Below-ground sink limitation alters growth and carbon balance of Eucalyptus seedlings

### 2015-10-01 manuscript/manuscript_pve.Rmd
edits from MT

\
Alternatively, manipulating plant tissue C sinks is often used to investigate the correlation of A and growth. This is because metabolic signaling networks, relaying information on C and N status of different tissues, can regulate photosynthetic activity [@paul2001sink]. If sink inhibition of A occurs, a close coordination between declines in A and growth should be expected. Whether photosynthetic down regulation is evident in woody species has been tested through fruit removal, phloem girdling, and low temperatures at high elevations. In these studies, down regulation of A was frequently correlated to carbohydrate accumulation resulting from reduced tissue sink strength  [@iglesias2002regulation;@hoch2002altitudinal; @urban2007girdling; @haouari2013fruit]. However, reductions in A were also attributed to biochemical limitations prior to carbohydrate accumulation [@nebauer2011photosynthesis], irreversible photo-oxidative damage [@duan2008photosynthetic], and stomatal limitation [@li2005photosynthesis]. These mixed results are not surprising as we still know little about the pathways in which plants achieve balance between assimilation, storage, and growth across temporal scales [@smith2007coordination]. As these manipulations likely impact source as well as sink activity simultaneously, affect water transport, are very extreme, or are specific to the occurence large fruiting sinks, they tell us little about source-sink coordination in typical growing conditions for woody species.

\
An alternative experimental approach is to reduce belowground C sink strength in tree seedlings by manipulating rooting volume, by varying the container size. The advantage of this approach is that it allows a large range of treatment levels, can be easily compared to naturally planted seedlings and may mimic natural conditions as seedlings compete for space or reach bedrock. Seedlings undergo many physiological and morphological changes in response to rooting volume, including biomass partitioning, A, water relations, nutrient uptake and respiration [@nesmith1998effect;@poorter2012pot and references therein]. Inadequate rooting volume may decrease C sink strength by progressively restricting root growth in growing plants [@thomas1991root].  Container size studies frequently exhibit photosynthetic down-regulation, likely as a result of sink limitation [@arp1991effects; @mcconnaughay1991physical; @gunderson1994photosynthetic; @sage1994acclimation; @maina2002intra; @ronchi2006growth]. A meta-analysis by Poorter et al. [-@poorter2012pot] concluded that A is the process likely to be the strongest affected by pot size and may best explain the effects on biomass seen in the large number of studies where containers are used. This conclusion arises because plants grown in small containers are shown to accumulate leaf starch while having lower C exchange and assimilate export rates [@robbins1988effect]. However, evidence in support for a trade-off between C storage and growth in trees is, to date, inconclusive [@palacio2014does]. Based on these previous studies, using container size as a sink-strength manipulation can be used to empirically test the extent to which growth and A are coordinated.

@@

2). As the finite pool of rooting volume and soil nutrients will decline faster in trees growing in small containers, we expected reductions in partitioning to fine root mass to be magnified with decreasing container size.

3). Reduced sink strength was expected to lead to accumulation of leaf non-structural carbohydrates, and a resulting downregulation of A. We therefore expected a correlation between carbohydrate accumulation and photosynthetic capacity as a function of soil volume.

@@

## Growth and morphology metrics
Seedlings were planted in summer (January 21^st^ 2013) and stem height, diameter at 15 cm and leaf count were measured weekly thereafter. Once the growth rate of individual plants had significantly declined a full biomass harvest was completed abd the experiment ended (May 21^st^ 2013). Dry mass of leaves, stems, roots and total leaf area (LI-3100C Area Meter; LI-COR, Lincoln, NE, USA) were measured for each seedling. Mean individual leaf area for each harvested seedling was calculated by dividing total measured leaf area by total leaf count of only fully expanded leaves. Mean individual leaf area was then used to interpolate total seedling leaf area through time with weekly leaf counts. Root mass was collected by removing the roots system and passing soil from each container through a 1 mm sieve, washing, separating into fine and coarse roots (<2 mm and >2 mm diameter, respectively) and then drying to a constant mass. Roots from the free seedlings were collected by excavating each 1 m^2^ subplot to the hard layer and keeping only roots within the subplot. 25 g fresh weight subsamples of washed fine roots were analyzed, using WhinoRhizo software (Regent Instruments Inc., Quebec, QC, Canada), for specific root length (SRL, m g^-1^).

@@

\
(1)
$$P_i = L\left(\frac{C_{day'i}~\sigma_s}{\epsilon_c}\right) - R$$
where L is total plant leaf area (m^2^), C~day,i~ is the predicted daily carbon assimilation (g C g mass^-1^ d^-1^), $\sigma$~s~ is a self shading parameter, $\epsilon$~c~ is a biomass conversion efficiency parameter and R is the mass based total respiration of all tissue components (g C d^-1^). Total respiration was calculated as

\
(2)
$$R = \Sigma(R_c~M_c)$$
where R~c~ is publsihed or local tissue respiration rates of fine roots, coarse roots or stems (Table S1, g C g mass ^-1^ d^-1^) and M~c~ is the standing biomass of each component (g). R~leaf~ is represented in the calculation of C~day~ (described below). The change in individual component biomass (M~c~), here solved on a daily time step, is given by

@@

##Data analysis
Differences in experimental parameters with soil volume were analysed by mixed-effects models in R with individual containers as random effects and soil volume treatment as a categorical fixed effect iwth seven levels. Tukey's post-hoc tests were performed in conjunction with ANOVA to determine which specific paired comparisons among soil volume treatments were different. A linear mixed effect model of A~max~ and leaf chemistry was performed using the 'nlme' package [@nlme] in R. Explained variance (R^2^) of mixed models were computed as in  [@nakagawa2013general]. Tests of allometric relationships between log-transformed biomass components were implemented using standardized major axis regression in the 'smatr' package in R [@warton2012smatr]. Results were considered significant at P < 0.05.

@@

## Growth and morphology

Plant height, diameter, and leaf area diverged between container volumes soon after start of the experiment (Figure `r allomfig`). First, seedling leaf area significantly diverged between soil volumes (P < 0.026) during the 5^th^ week of the experiment. Following this period both height (8^th^) week) and then diameter (9^th^ week) significantly deviated across soil volumes (P < 0.002 & 0.001, respectively). The large reductions in height gain and total leaf area in smaller compared to larger containers continued throughout the experiment. In this field study, colder temperatures and reductions in total PPFD per day (Figure `r airvarsfig`) likely slowed the growth of seedlings in the final weeks of the experiment (Figure `r allomfig`). Seedlings maintained diameter growth throughout the experiment, although marginal with smaller soil volumes in the final month. Final seedling height significantly increased with increasing soil volume (P < 0.001). Increases in both final stem diameter (P < 0.001) and cumulative leaf area (both P < 0.001) were found with increasing soil volume and these differences were driven mainly by the largest container and the free seedling treatments.

\
Total seedling biomass at harvest was significantly different across container volumes (P < 0.001) and between container treatments and free seedlings (P < 0.001, Table 1). On average, harvested biomass of free seedlings was 84% higher than seedlings in containers. We analyzed the relationship between biomass growth and soil volume and found an increase of 34 % with a doubling of container volume, consistent with the meta-analysis of Poorter et al. [-@poorter2012pot]. Additionally, plant biomass was highly correlated with total leaf area across all treatments (R^2^ = 0.97, P < 0.001). Differences in biomass partitioning to leaves, stems, and roots were not different across soil volumes after variation in seedling biomass across treatments was factored in the analysis (Figure `r allocfig`a,b). Across all treatments, the final harvest root:shoot biomass ratio was conserved in these seedlings which exhibited a slightly higher shoot than root mass ($\overline{x}$ = 0.904, 95% CI = [0.846,1.119]) and a near identical ratio of leaf to fine root mass (Figure `r allocfig`c).

@@

\
Next, we performed a series of model simulations to test possible C allocation scenarios to account for this excess pool of C. Testing the sensitivity of the model to treatment-specific LMF from the final harvest (M~2~) improved model predictions of seedling C mass but still overestimated seedling total C by 32&pm;11.1 % (Figure `r massmodelfig2`a). Using harvest values of LMF, however, does not capture the increase in senescence of seedlings in small containers (Figure `r allomfig`c). Thus, the 7-61 % reduction in harvest avlues of LMF from large to small soil volumes compared to the optimized LMF (M~0~) represents an underestimation of realized leaf C allocation. Increases of 50 % in non-leaf tissue respiration (M~3~) improved biomass estimates slightly but overestimated mass C by an average of 46&pm;9.3 % in seedlings with soil volume restriction (Figure `r massmodelfig2`b). With M~3~, non-tissue respiration rates would need to be increased by ca. 250% in order for mass balance to be achieved.

# Discussion
This study utilized a simple but novel field design to manipulate belowground sink limitation and physically restrict *Eucalyptus tereticornis* seedling biomass production. We addressed questions regarding the coordination of A and growth by complementing empirical results with a carbon balance model. We found that reductions in leaf A across container sizes were not sufficient to explain observed reductions in total plant biomass production. We thus encourage the utilization of mass balance approaches and provide direction for future studies when testing factors that control plant growth under environmental change.

@@

\
Although the soil N % declined evenly across all treatments, leaf N was lowest in the smallest containers, suggesting sink limitation was the greatest in these containers. This makes sense as small containers may result in a reduction of total N uptake, either from physical root restriction or decreased supply, which will affect growth, Rubisco limitation, sugar metabolism, and carbohydrate partitioning between source and sink tissues [@stitt1991rising; @hermans2006plants]. Mycorrhizal colonization could also have been affected in containers. Whether unrestricted mycorrhizal recruitment facilitated the increase in leaf N uptake in free seedlings is unknown. Regardless of whether sink limitation occurred from root or mycorrhizal restriction, the stark contrast in leaf N between free seedlings and seedlings in containers was significantly correlated to reductions in A.

@@

## Biomass partitioning under sink limitation
As biomass partitioning is likely controlled by the source and sink strength of all organs [@poorter2012biomass], it was important to determine which tissue components were most affected by the container size treatments. It was necessary to distinguish if growth was affected beyond ontogenetic constraints, by correcting for size, as biomass distribution is strongly size-dependent [@gould1966allometry; @lleonart2000removing]. In this study, there was no significant difference in root, leaf, or stem biomass partitioning with reduced soil volume compared to free seedlings, outside of ontogenetic drift. This is a surprising result as shifts in allocation have been noted specifically for nutrient limitation  [@mcconnaughay1999biomass, and references therein]. Surprisingly, a constant ratio of fine root mass to leaf mass was observed across all treatments suggesting a functional partitioning response to optimize resource gain did not occur.

\
As partitioning to fine roots did not change this provides evidence against an optimal foraging strategy for seedlings in containers. This could be because lateral root development is affected by inanimate obstacles and avoiding growth towards container walls could improve the efficiency of resource allocation [@falik2005root]. Also, the sensitivity of roots to their own exudates near obstructions may be used to adjust growth before stressful conditions occur [@semchenko2008foraging]. Alternatively, physical restriction of root proliferation could have impacted root development and morphology prior to shifts in mass partitioning. Here, increases in root length where detected in several of the soil volume treatments. This is not surprising as plants in containers have been shown to have different root morphology to field grown plants [@nesmith1998effect]. The poor soil quality used in our experiment and root restriction, however, likely decreased the capacity of this morphological response to increase N uptake.

## Do reductions in photosynthesis explain reductions in seedling growth?
Our growth model used a simple but conventional approach to drive seedling growth with fluctuations in A while treating C use efficiency, respiration and C allocation as fixed processes. Contrary to expectation, the model consistently overestimated seedling growth in containers when parameterized with an optimized LMF for free seedlings. Although reductions in A~max~ and biomass were strongly correlated among treatments, as hypothesized by Poorter et al. [-@poorter2012pot], we provide evidence that links between sink limitation and A do not necessarily imply the same coordination between reduced A and growth. These findings are important as this model reflects classical approaches in tree growth and production modelling that are driven by inputs of C assimilation and processes such as respiration are considered proportional to biomass [@le2001carbon]. Our results indicate a need to evaluate the use of fixed processes in models which distinguish the fate of assimilate C within a plant. Doing so will provide valuable input to future models as assimilate allocation is a key component in carbon-balance driven  plant growth models, yet C partitioning remains a key weakness  [@lacointe2000carbon]. To address this issue we utilized the flexibility of this model to test plausible fates of the extra pool of non-biomass C unaccounted for with mass balance. Similar to Lohier et al. [-@lohier2014explaining] we manipulated processes contributing to seedling C mass balance, including changes to leaf C allocation or non-leaf tissue respiration, to quantitatively test their respective influences on model predictions.

@@

\
**Figure 5**. Photosynthetic capacity, on a leaf mass basis, as a function of accumulation of leaf starch (a) and leaf nitrogen content without TNC (b).  Colors represent bins levels (n = 5) of both leaf starch and nitrogen grouped from low to high .  Lines represents predictions, for each bin level, from the linear mixed effects model equation of A~mass~ as a function of starch and nitrogen. The marginal R^2^ (fixed effects only) was 0.37 and the conditional R^2^ (fixed and random effects) was 0.48 for the complete model.

>>>>

\
Alternatively, manipulating plant tissue C sinks is often used to investigate the correlation of A and growth. This is because metabolic signaling networks, relaying information on C and N status of different tissues, can regulate photosynthetic activity [@paul2001sink]. If sink inhibition of A occurs, a close coordination between declines in A and growth should be expected. Whether photosynthetic down regulation is evident in woody species has been tested through fruit removal, phloem girdling, and low temperatures at high elevations. In these studies, down regulation of A was frequently correlated to carbohydrate accumulation resulting from reduced tissue sink strength  [@iglesias2002regulation;@hoch2002altitudinal; @urban2007girdling; @haouari2013fruit]. However, reductions in A were also attributed to biochemical limitations prior to carbohydrate accumulation [@nebauer2011photosynthesis], irreversible photo-oxidative damage [@duan2008photosynthetic], and stomatal limitation [@li2005photosynthesis]. These mixed results are not surprising as we still know little about the balance between assimilation, storage, and growth across temporal scales in plants [@smith2007coordination]. As these manipulations likely impact source as well as sink activity simultaneously, affect water transport, are very extreme, or are specific to the occurence large fruiting sinks, they tell us little about source-sink coordination in typical growing conditions for woody species.

\
An alternative experimental approach is to reduce belowground C sink strength in tree seedlings by manipulating rooting volume, by varying the container size [@arp1991effects; @nesmith1998effect; @poorter2012pot]. Possible advantages of this approach are that it allows a large range of treatment levels, can be easily compared to naturally planted seedlings and may mimic natural conditions as seedlings compete for space or reach bedrock. Seedlings undergo many physiological and morphological changes in response to rooting volume, including biomass partitioning, A, water relations, nutrient uptake and respiration [@nesmith1998effect;@poorter2012pot and references therein]. Inadequate rooting volume may decrease C sink strength by progressively restricting root growth in growing plants [@thomas1991root].  Container size studies frequently exhibit photosynthetic down-regulation, likely as a result of sink limitation [@arp1991effects; @mcconnaughay1991physical; @gunderson1994photosynthetic; @sage1994acclimation; @maina2002intra; @ronchi2006growth]. A meta-analysis by Poorter et al. [-@poorter2012pot] concluded that A is the process likely to be the strongest affected by pot size and may best explain the effects on biomass seen in the large number of studies where containers are used. This conclusion arises because plants grown in small containers are shown to accumulate leaf starch while having lower C exchange and assimilate export rates [@robbins1988effect]. However, evidence in support for a trade-off between C storage and growth in trees is, to date, inconclusive [@palacio2014does]. Based on these previous studies, using container size as a sink-strength manipulation can be used to empirically test the extent to which growth and A are coordinated.

@@

2). As the finite pool of rooting volume and soil nutrients will decline faster in trees growing in small containers, we expected reductions in partitioning to fine root mass relative to tree size with decreasing container size.

3). Reduced sink strength was expected to lead to accumulation of leaf non-structural carbohydrates, and a resulting down regulation of A. We therefore expected a correlation between carbohydrate accumulation and photosynthetic capacity as a function of soil volume.

@@

## Growth and morphology metrics
Seedlings were planted in summer (January 21^st^ 2013) and stem height, diameter at 15 cm and leaf count were measured weekly thereafter. Once the growth rate of individual plants had significantly declined a full biomass harvest was completed abd the experiment ended (May 21^st^ 2013). Dry mass of leaves, stems, roots and total leaf area (LI-3100C Area Meter; LI-COR, Lincoln, NE, USA) were measured for each seedling. Mean individual leaf area for each harvested seedling was calculated by dividing total measured leaf area by total leaf count of only fully expanded leaves. Mean individual leaf area was then used to interpolate total seedling leaf area through time with weekly leaf counts. Root mass was collected by removing the roots system and passing soil from each container through a 1 mm sieve, washing, separating into fine and coarse roots (<2 mm and >2 mm diameter, respectively) and then drying to a constant mass. Roots of seedlings in containers were not considered pot bound, as clusters of roots along the soil-container interface were not observed. Roots from the free seedlings were collected by excavating each 1 m^2^ subplot to the hard layer and keeping only roots within the subplot. For each seedling, a 25 g subsample of washed fine roots was analyzed for root length using WhinoRhizo software (Regent Instruments Inc., Quebec, QC, Canada). Here, specific root length (SRL) is reported as the root length divided by the dry mass of each subsample (m g^-1^).

@@

\
(1)
$$P_i = L\left(\frac{C_{day'i}~\sigma_s}{\epsilon_c}\right) - R$$
where L is total plant leaf area (m^2^), C~day,i~ is the predicted daily carbon assimilation (g C g mass^-1^ d^-1^), $\sigma$~s~ is a self shading parameter (explained below), $\epsilon$~c~ is a biomass conversion efficiency parameter (g C g mass^-1^) and R is the mass based total respiration of all tissue components (g C d^-1^). Total respiration was calculated as

\
(2)
$$R = \Sigma(R_c~M_c)$$
where R~c~ is publsihed or local tissue respiration rates of fine roots, coarse roots or stems (Table S1, g C g mass^-1^ d^-1^) and M~c~ is the standing biomass of each component (g). R~leaf~ is represented in the calculation of C~day~ (described below). The change in individual component biomass (M~c~), here solved on a daily time step, is given by

@@

##Data analysis
Differences in experimental parameters with soil volume were analysed by mixed-effects models in R with individual containers and experimental blocks as random effects and soil volume treatment as a categorical fixed effect with seven levels. Tukey's post-hoc tests were performed in conjunction with ANOVA to determine which specific paired comparisons among soil volume treatments were different. A linear mixed effect model of A~max~ and leaf chemistry was performed using the 'nlme' package [@nlme] in R. Explained variance (R^2^) of mixed models were computed as in  [@nakagawa2013general]. Tests of allometric relationships between log-transformed biomass components were implemented using standardized major axis regression in the 'smatr' package in R [@warton2012smatr]. Results were considered significant at P < 0.05.

@@

## Growth and morphology
Plant height, diameter, and leaf area diverged between container volumes soon after start of the experiment (Figure `r allomfig`). First, seedling leaf area significantly diverged between soil volumes (P < 0.029) during the 5^th^ week of the experiment. Following this period both height (8^th^) week) and then diameter (9^th^ week) significantly deviated across soil volumes (P < 0.045 & 0.035, respectively). The large reductions in height gain and total leaf area in smaller compared to larger containers continued throughout the experiment. In this field study, colder temperatures and reductions in total PPFD per day (Figure `r airvarsfig`) likely slowed the growth of seedlings in the final weeks of the experiment (Figure `r allomfig`). Seedlings maintained diameter growth throughout the experiment, although marginal with smaller soil volumes in the final month. Final seedling height significantly increased with increasing soil volume (P < 0.001). Increases in both final stem diameter (P < 0.001) and cumulative leaf area (both P < 0.001) were found with increasing soil volume and these differences were driven mainly by the largest container and the free seedling treatments.

\
Total seedling biomass at harvest was significantly different across container volumes (P < 0.001) and between container treatments and free seedlings (P < 0.001, Table 1). On average, harvested biomass of free seedlings was 84% higher than seedlings in containers. We analyzed the relationship between biomass growth and soil volume and found an increase of 34 % with a doubling of container volume, consistent with the meta-analysis of Poorter et al. [-@poorter2012pot]. Additionally, plant biomass was postively correlated with total leaf area across all treatments (R^2^ = 0.97, P < 0.001). Differences in biomass partitioning to leaves, stems, and roots were not different across soil volumes after variation in seedling biomass across treatments was factored in the analysis (Figure `r allocfig`a,b). Across all treatments, the final harvest root:shoot biomass ratio was conserved in these seedlings which exhibited a slightly higher shoot than root mass ($\overline{x}$ = 0.904, 95% CI = [0.846,1.119]) and a near identical ratio of leaf to fine root mass (Figure `r allocfig`c).

@@

\
Next, we performed a series of model simulations to test possible C allocation scenarios to account for the over predictions of seedling C mass. Testing the sensitivity of the model to treatment-specific LMF from the final harvest (M~2~) improved model predictions of seedling C mass but still overestimated seedling total C by 32&pm;11.1 % (Figure `r massmodelfig2`a). Using harvest values of LMF, however, does not capture the increase in leaf turnover of seedlings in small containers (Figure `r allomfig`c). Thus, the 7-61 % reduction in harvest avlues of LMF from large to small soil volumes compared to the optimized LMF (M~0~) represents an underestimation of realized leaf C allocation. Increases of 50 % in non-leaf tissue respiration (M~3~) improved biomass estimates slightly but overestimated mass C by an average of 46&pm;9.3 % in seedlings with soil volume restriction (Figure `r massmodelfig2`b). With M~3~, non-tissue respiration rates would need to be increased by ca. 250% in order for mass balance to be achieved.

# Discussion
This study utilized a simple but novel field design to manipulate belowground sink limitation and physically restrict *Eucalyptus tereticornis* seedling biomass production. We addressed questions regarding the coordination of A and growth by complementing empirical results with a C balance model. We found that reductions in leaf A across container sizes, when integrated across the 120 day experiment, were alone insufficient to account for observed reductions in total plant biomass production.

We thus encourage the utilization of mass balance approaches and provide direction for future studies when testing factors that control plant growth under environmental change.

@@

\
Here, reductions in A were postively correlated with decreases in leaf N and leaf N was considerably reduced for seedlings in containers. Physical root restriction or decreased supply as been shown cause reductions in N uptake (citations), which could explain the observed reductions in leaf N in the smallest containers. Root N at the end of the experiment was on average higher in free seedlings but not consistently higher than every container volume treatment. Unrestricted mycorrhizal recruitment could have instead faciliated the increases in leaf N in free seedlings, but this effect is unknown. Combined with the fact that soil N % declined evenly across all treatments, there was no clear mechanism present to identify changes in root N uptake between free seedlings and seedlings in containers. In these already low quality soils, it is possible that seedlings in containers simply grew into increasing N limitation which negatively affected belowground sink strength. Although no clear feedback could be determined between the available soil N pool and decreases in leaf N, the affects of belowground sink limitation on A of seedlings in containers was evidents throughout the experiment.

@@

## Biomass partitioning under sink limitation
As biomass partitioning is likely controlled by the source and sink strength of all organs [@poorter2012biomass], it was important to determine which tissue components were most affected by the container size treatments. It was necessary to distinguish if growth was affected beyond ontogenetic constraints, by correcting for size, as biomass distribution is strongly size-dependent [@gould1966allometry; @lleonart2000removing]. In this study, there was no significant difference in root, leaf, or stem biomass partitioning with reduced soil volume compared to free seedlings, outside of ontogenetic drift (Figure `r allocfig`a,b). This is a surprising result as shifts in allocation have been noted specifically for nutrient limitation  [@mcconnaughay1999biomass, and references therein]. Surprisingly, a constant ratio of fine root mass to leaf mass was observed across all treatments suggesting a functional partitioning response to optimize resource gain did not occur.

\
As partitioning to fine roots did not change this provides evidence against an optimal foraging strategy for seedlings in containers. This could be because lateral root development is affected by inanimate obstacles and avoiding growth towards container walls could improve the efficiency of resource allocation [@falik2005root]. Also, the sensitivity of roots to their own exudates near obstructions may prevent further growth [@semchenko2008foraging]. Alternatively, physical restriction of root proliferation could have impacted root development and morphology prior to shifts in mass partitioning. Here, increases in root length where detected in several of the soil volume treatments. This is not surprising as plants in containers have been shown to have different root morphology to field grown plants [@nesmith1998effect]. The poor soil quality used in our experiment and root restriction, however, likely decreased the capacity of this morphological response to increase N uptake.

## Do reductions in photosynthesis explain reductions in seedling growth?
Our conventional growth model used a simple approach to drive seedling growth with fluctuations in A while treating C use efficiency, respiration and C allocation as fixed processes. Contrary to expectation, the model consistently overestimated seedling growth in containers when parameterized with an optimized LMF for free seedlings. Although reductions in A~max~ and biomass were strongly correlated among treatments, as hypothesized by Poorter et al. [-@poorter2012pot], we provide evidence that links between sink limitation and A do not necessarily imply the same coordination between reduced A and growth. These findings are important as this model reflects classical approaches in tree growth and production modelling that are driven by inputs of C assimilation and processes such as respiration are considered proportional to biomass [@le2001carbon]. Our results indicate a need to evaluate the use of fixed processes in models which distinguish the fate of assimilate C within a plant. Doing so will provide valuable input to future models as assimilate allocation is a key component in carbon-balance driven  plant growth models, yet C partitioning remains a key weakness  [@lacointe2000carbon]. To address this issue we utilized the flexibility of this model to test plausible fates of the extra pool of non-biomass C unaccounted for with mass balance. Similar to Lohier et al. [-@lohier2014explaining] we manipulated processes contributing to seedling C mass balance, including changes to leaf C allocation or non-leaf tissue respiration, to quantitatively test their respective influences on model predictions.

@@

\
**Figure 5**. Photosynthetic capacity, on a leaf mass basis, as a function of accumulation of leaf starch (a) and leaf nitrogen content without TNC (b).  Colors represent bins levels (n = 5) of both leaf starch and nitrogen grouped from low to high.  Lines represents predictions, for each bin level, from the linear mixed effects model equation of A~mass~ as a function of starch and nitrogen. The marginal R^2^ (fixed effects only) was 0.37 and the conditional R^2^ (fixed and random effects) was 0.48 for the complete model.
